---
title: "Closing Celebration Challenge"
author: "sg"
date: "3/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages}
library(tidyverse)
library(viridis)
library(lubridate)
library(scales)
library(sf)
library(tmap)
library(kableExtra)
library(ggraph)
library(calendR)
library(tidygraph)
library(ggmosaic)
```

The data is coming from Chicago's data portal.  I have elected to combine a number of datasets under the Traffic Crashes umbrella (People, Crashes, Vehicles).

```{r, echo=FALSE}
path <- '/Users/SG/Documents/Programming/dac2022/closing_celebration/'

crashes <- janitor::clean_names(read_csv('Traffic_Crashes_-_Crashes.csv'))
people <- janitor::clean_names(read_csv(paste(path,'Traffic_Crashes_-_People.csv', sep = '')))
vehicles <- janitor::clean_names(read_csv(paste(path, 'Traffic_Crashes_-_Vehicles.csv', sep = '')))


#vehicles <- jsonlite::read_json('https://data.cityofchicago.org/resource/68nd-jvt3.json?$limit=200000', simplifyVector = TRUE)

#crashes <- jsonlite::read_json('https://data.cityofchicago.org/resource/85ca-t3if.json?$limit=200000', simplifyVector = TRUE)

#people <- jsonlite::read_json('https://data.cityofchicago.org/resource/u6pd-qa9d.json?$limit=200000', simplifyVector = TRUE)

`%!in%` <- Negate(`%in%`)
count_summary <- function(data){
  data %>%
  filter(maneuver %!in% c('UNKNOWN/NA', 'NA'),
         area_hit %!in% c('UNKNOWN', 'NA')) %>%
  drop_na(maneuver, area_hit) %>%
  group_by(maneuver, area_hit) %>%
  summarise(n = n())
}
```

```{r}
flag_colors <- c('#e63946', '#f1faee', '#a8dadc', '#457b9d', '#1d3557')

grey_red <- c('#2b2d42', '#8d99ae', '#edf2f4', '#ef233c', '#d90429')

bg_color <- '#f8f9fa'

bg_photo <- '/Users/SG/Documents/Programming/dac2022/closing_celebration/pedro-lastra-Nyvq2juw4_o-unsplash.jpg'

bg_photo2 <- '/Users/SG/Documents/Programming/dac2022/closing_celebration/neal-kharawala-XXA8PTuLD1Y-unsplash.jpg'

theme_cel <- function(base_size = 11,
                      base_family = 'serif'){
    theme(
      rect = element_rect(fill = bg_color, colour = NA, linetype = 0),
      line = element_line(color = grey_red[1]),
      
      axis.ticks = element_line(),
      axis.ticks.y = element_blank(),
      axis.line = element_line(linetype = 1, colour = grey_red[1]),
      
      legend.key = element_rect(linetype = 0, fill = bg_color),
      
      panel.background = element_rect(linetype = 0, fill = bg_color),
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = bg_color, colour = NA, linetype = 0),
      plot.background = element_rect(fill = bg_color, colour = NA))
}
```

Cleaning the data

```{r, echo=FALSE}
# Some preliminary cleaning and joining Vehicles + Crashes
vehicles <- vehicles %>%
  select(id = crash_record_id,
         date = crash_date,
         make,
         model,
         type = vehicle_type,
         maneuver,
         passengers = num_passengers,
         area_hit = first_contact_point) %>%
  mutate(date = mdy_hms(date),
         passengers = replace_na(passengers, 0))

people <- people %>%
  select(id = crash_record_id,
         action = driver_action,
         injury = injury_classification,
         airbag = airbag_deployed,
         zipcode,
         sex,
         age)  %>%
  mutate(age_group = case_when(
    age < 21 ~'Under 21',
    age >= 21 & age <= 44 ~'21-44',
    age > 44 & age < 65 ~'45-64',
    age >= 65 ~'65 and over',
    TRUE ~'NA'),
    airbag = if_else(str_detect(airbag, 'DEPLOYED'), TRUE, FALSE),
    airbag = replace_na(airbag, FALSE),
    is_injured = if_else(str_detect(injury, 'NO INDICATION'), FALSE, TRUE))

crashes <- crashes %>%
  select(id = crash_record_id,
         street = street_name,
         weather = weather_condition,
         roadtype = trafficway_type,
         speed = posted_speed_limit,
         road_cond = roadway_surface_cond,
         crash_type,
         damage,
         latitude,
         longitude,
         location) %>%
  mutate(is_severe = if_else(str_detect(crash_type, 'TOW'), TRUE, FALSE),
         four_way = if_else(roadtype == 'FOUR WAY', TRUE, FALSE),
         bad_weather = if_else(weather %in% c('RAIN', 'FREEZING RAIN/DRIZZLE', 'SLEET/HAIL', 'SNOW'), TRUE, FALSE),
         expensive = if_else(damage %in% 'OVER $1,500', TRUE, FALSE),
         wet_road = if_else(str_detect(road_cond, 'SNOW | WET | ICE'), TRUE, FALSE),
         year = year(date))

traffic <- left_join(crashes, vehicles, by = 'id')
```


```{r}
crash_dummy <- traffic %>%
  mutate(is_severe = if_else(str_detect(crash_type, 'TOW'), TRUE, FALSE),
         four_way = if_else(roadtype == 'FOUR WAY', TRUE, FALSE),
         bad_weather = if_else(weather %in% c('RAIN', 'FREEZING RAIN/DRIZZLE', 'SLEET/HAIL', 'SNOW'), TRUE, FALSE),
         expensive = if_else(damage %in% 'OVER $1,500', TRUE, FALSE))

crash_mod <- glm(is_severe~ bad_weather + four_way, data = crash_dummy, binomial(link = 'probit'))

summary(crash_mod)

crash_pass <- glm(is_severe~ passengers, data = crash_dummy, binomial(link = 'probit'))

money_rain <- glm(expensive~ bad_weather, data = crash_dummy, binomial(link = 'probit'))

summary(crash_pass)
summary(money_rain)

severe_wet <- glm(expensivee~  bad_weather * wet_road, data = traffic, binomial(link = 'probit'))

summary(severe_wet)

plot(severe_wet)

```

Some exploratory plots.  The danger with some of these simple counting plots is I am likely counting the most popular cars, as opposed to the cars most likely to get into an accident.  It would be nice if I could find some IL or Chicago data about vehicle market share, but I haven't been able to find a suitable dataset.

``` {r}
traffic %>%
  filter(!is.na(make),
         make != 'UNKNOWN') %>%
  mutate('make' = fct_lump(make, n = 20)) %>%
  ggplot() +
  geom_histogram(aes(fct_rev(make)), stat = 'count', fill = flag_colors[5]) +
  coord_flip() +
  theme_cel() +
  xlab('') +
  ylab('') +
  scale_y_continuous(expand = c(0,0)) +
  labs(title = 'Number of Accidents by Car Make',
       subtitle = 'in the past 4 days Chicago, IL')

traffic %>%
  filter(!is.na(model),
         make != 'UNKNOWN',
         model != 'OTHER (EXPLAIN IN NARRATIVE)',
         make != 'MOTORIZED') %>%
  group_by(model, make) %>%
  summarise('count' = n()) %>%
  arrange(desc(count)) %>%
  head(12) %>%
  ggplot() +
  geom_col(aes(x = fct_rev(fct_reorder(model, count)), y = count, fill = make), color = 'gray') +
  scale_fill_manual(values = flag_colors) +
  labs(title = 'What make and model cars are getting in the most accidents?',
       subtitle = '...or maybe simply which are popular',
       fill = 'Make') +
  ylab('Number of Accidents') +
  xlab('Model') +
  theme_cel()
  

# The x axis here is fucking brutal
traffic %>%
  group_by(time = hour(date)) %>%
  summarise(count = n()) %>%
  ggplot() +
  geom_col(aes(x = time, y = count))

traffic %>%
  ggplot() +
  geom_histogram(aes(maneuver), stat = 'count') +
  coord_flip()


traffic %>%
  ggplot() +
  geom_histogram(aes(occupants), stat = 'count')

traffic %>%
  filter(maneuver %!in% c('UNKNOWN/NA', 'OTHER'),
         area_hit %!in% c('UNKNOWN', 'OTHER')) %>%
  ggplot() +
  geom_jitter(aes(x = maneuver, y = area_hit)) +
  scale_x_discrete(guide = guide_axis(angle = 90))
```


```{r}


chi_sf %>%
ggplot() +
  geom_sf() +
  geom_sf(data = traffic_sf, aes(color = count), alpha = 0.4)
```

```{r}
setwd('/Users/SG/Documents/Programming/dac2022/L9_Upload')
chi_sf <- st_read("chicago_tracts_2010.shp")
chi_sf <- st_transform(chi_sf, 4326)
traffic_sf <- traffic %>% drop_na(c(latitude, longitude)) %>%
  st_as_sf(coords = c('longitude', 'latitude'), crs = 4326)

traffic_sf <- st_join(
  traffic_sf,
  chi_sf, 
  join = st_within
  )

traffic_sf$geometry.1 <- NULL
traffic_sf <- traffic_sf %>%
  group_by(geoid) %>%
  summarise(count = n())

traffic_sf_test <- st_join(
  chi_sf,
  traffic_sf
  )

tmap_mode('view')
tm_shape(traffic_sf, name="Chicago Accidents") + tm_bubbles(size = 'count', col = grey_red[5], scale = 1/3) +
  tm_basemap('Esri.WorldStreetMap')
```


Creating a contingency table for maneuver/area_hit

```{r}
traffic_kable <- traffic %>%
  filter(maneuver %!in% c('UNKNOWN/NA', 'NA'),
         area_hit %!in% c('UNKNOWN', 'NA')) %>%
  drop_na(maneuver, area_hit) %>%
  group_by(maneuver, area_hit) %>%
  summarise(n = n()) %>%
  mutate(prop = n/sum(n) * 100) %>%
  pivot_wider(id_cols = maneuver,
              names_from = c(area_hit),
              values_from = c(prop)) %>%
  mutate_at(c(1:14), ~replace(., is.na(.), 0)) %>%
  mutate_if(is.numeric, round) %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = 'Cambria')
  
traffic_kable
```

Consider Correlation Matrix (LOL NO)
```{r}
library(Hmisc)
library(corrplot)
traffic_corr <- traffic %>%
  filter(maneuver %!in% c('UNKNOWN/NA', 'NA'),
         area_hit %!in% c('UNKNOWN', 'NA')) %>%
  drop_na(maneuver, area_hit) %>%
  group_by(maneuver, area_hit) %>%
  summarise(n = n())

library(corrplot)
traffic_corr2 <- traffic %>%
  filter(maneuver %!in% c('UNKNOWN/NA', 'NA'),
         area_hit %!in% c('UNKNOWN', 'NA')) %>%
  drop_na(maneuver, area_hit) %>%
  group_by(maneuver, area_hit) %>%
  summarise(n = n()) %>%
  pivot_wider(id_cols = maneuver,
              names_from = c(area_hit),
              values_from = c(n)) %>%
  mutate_at(c(1:14), ~replace(., is.na(.), 0))

traffic_corr3 <- matrix(rep(FALSE, 286), nrow = 22)

traffic_corr2 <- as.matrix(traffic_corr2)
names <- traffic_corr2[,1]
colnames <- colnames(traffic_corr2)
rownames(traffic_corr2) <- names
traffic_corr2 <- traffic_corr2[,-1]
traffic_corr3[] <- as.numeric(traffic_corr2)
rownames(traffic_corr3) <- names
colnames(traffic_corr3) <- colnames


corrplot(cor(traffic_corr3),
         method = 'shade',
         type = 'full',
         diag = FALSE,
         tl.col = 'black',
         bg = 'white'
         )
```


Let's do use ggraph to make a network graph
```{r}
traffic_connect_prop <- traffic %>%
  mutate(area_hit = replace(area_hit, area_hit == 'OTHER', 'OTHER AREA')) %>%
  filter(year == 2022) %>%
  count_summary() %>%
  mutate(percent = n/sum(n) * 100) %>%
  rename(crashes = n)

traffic_connect_prop %>%
  ggraph(layout = "linear", circular = TRUE) +
  geom_edge_arc(aes(edge_alpha = percent, edge_width = percent), color = flag_colors[1], fold = TRUE, lineend = 'round') +
  geom_node_text(aes(filter = name %in% traffic_connect_prop$area_hit, label = name), size = 3) +
  geom_node_label(aes(filter = name %in% traffic_connect_prop$maneuver, label = name), size = 3, repel = T, fill = flag_colors[2]) +
  theme_cel() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank()) +
  labs(title = '% of Accidents Resulting in First Point of Contact, Given a Maneuver',
       subtitle = 'Labels are Maneuvers')



traffic_connect_count <- traffic %>%
  mutate(area_hit = replace(area_hit, area_hit == 'OTHER', 'OTHER AREA')) %>%
  count_summary() %>%
  rename(crashes = n)


traffic_connect_count %>%
  ggraph(layout = "linear", circular = TRUE) +
  geom_edge_arc(aes(edge_alpha = crashes, edge_width = crashes), color = 'darkred') +
  geom_node_text(aes(filter = name %in% traffic_connect_count$area_hit, label = name), size = 3, col = 'navy') +
  geom_node_text(aes(filter = name %in% traffic_connect_count$maneuver, label = name), size = 3, repel = T, col = 'black') +
  theme_graph()
```

Let's try a calendar heatmap
```{r}
scale_fn <- function(x) { x / sqrt(sum(x^2)) }

date_traffic <- traffic %>%
  filter(year(date) == 2022) %>%
  group_by(date = as.Date(date)) %>%
  summarise(accidents = n()) %>%
  mutate(scale_acc = scale_fn(accidents))


if (unique(year(date_traffic$date)) == year(Sys.Date())){
    
    days <- rep(min(date_traffic$scale_acc) - 0.05, 365)
    days[1:nrow(date_traffic)] <- date_traffic$scale_acc
} else {
    days <- date_traffic$scale_acc
}

calendR(year = as.numeric(unique(format(date_traffic$date, format = '%Y'))),
        special.days = days,
        gradient = TRUE,
        low.col = '#edf2f4',
        special.col = '#e63946',
        bg.img = bg_photo,
        title = paste('Heatmap of Traffic Accidents Year ', unique(year(date_traffic$date))),
        title.size = 20,
        title.col = grey_red[1],
        lty = 1,
        lwd = 0.1,
        col = grey_red[1],
        days.col = grey_red[1],
        day.size = 3,
        subtitle = 'Chicago, IL | Photo: Pedro Lastra',
        subtitle.col = grey_red[1],
        weeknames = c("Mo", "Tu",  
                      "We", "Th",
                      "Fr", "Sa",
                      "Su"),
        weeknames.col = 'black',
        weeknames.size = 4,
        font.family = 'sans')
```

Let's do something with the people
```{r}
people_range <- people %>%
  drop_na(sex, age) %>%
  filter(sex %!in% 'X') %>%
  mutate(age_range = case_when(age < 25 ~'under 25',
                               age < 45 ~'25-45',
                               age < 65 ~'45-65',
                               TRUE ~'over 65'))


people$age_group <- factor(people$age_group, levels = c('Under 21', '21-44', '45-64', '65 and over', NA))

people %>% 
  drop_na(age_group, sex) %>%
  filter(sex %!in% 'X') %>%
  ggplot() +
  geom_mosaic(aes(x = product(sex), fill = age_group)) +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) +
  labs(title = 'Who is getting into accidents in Chicago?') +
  ylab('Age Range') +
  scale_fill_manual(values = flag_colors[-2]) +
  theme_cel()

people %>%
  filter(action %!in% c('UNKNOWN', 'NONE', NA, 'OTHER'),
         !is.na(injury),
         injury %!in% c('NO INDICATION OF INJURY')) %>%
  ggplot(aes(x = sort(action), fill = injury)) +
  geom_bar(position = 'fill') +
  theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()) +
  xlab('') +
  ylab('') +
  labs(title = 'What Actions Lead to Serious Injury?',
       subtitle = 'If Injury and Action Recorded') +
  scale_fill_manual(values = flag_colors[-2]) +
  coord_flip()

people %>% filter(!is.na(zipcode)) %>%
  group_by(zipcode) %>%
  summarise(accidents = n()) %>%
  arrange(desc(accidents)) %>%
  head(10)


options(tigris_use_cache = TRUE)

chi_zip <- tigris::zctas(cb = TRUE,
                 starts_with = c('60'),
                 class = 'sf')

```

```{r}
people %>%
  drop_na(action) %>%
  filter(action %!in% c('NONE', 'UNKNOWN')) %>%
  ggplot() +
  geom_histogram(aes(action), stat = 'count') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

people %>%
  drop_na(action, age) %>%
  filter(action %!in% c('NONE', 'UNKNOWN', 0)) %>%
  ggplot() +
  geom_jitter(aes(x = age_group, y = action))

people %>%
  drop_na(sex, age_group) %>%
  filter(sex != 'X') %>% 
  ggplot() +
  geom_jitter(aes(sex, age_group))


choro$ggplot_scale = scale_fill_brewer()
choro$render()

geo <- geocode_zip(people$zipcode)
people_sf <- left_join(people, geo, by = 'zipcode') %>%
  drop_na(lat, lng)  %>%
  st_as_sf(coords = c('lng', 'lat'), crs = 4269)

chi_sf <- st_read(path)
chi_sf <- st_transform(chi_sf, 4269)

people_sf <- people_sf %>% st_join(chi_sf, join = st_within)

people_sf <- people_sf%>%
  drop_na(geoid) %>%
  group_by(geoid) %>%
  summarise(crash = n())

ggplot() +
  geom_sf(data = people_sf, aes(fill = crash, colour = crash))


people %>%
  group_by(zipcode) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

```

```{r}
people_mod <- people %>%
  filter(sex %!in% c('X', NA)) %>%
  mutate(is_male = if_else(sex == 'M', TRUE, FALSE))

model = glm(is_male~ age_group, data = people_mod, binomial(link = 'probit'))
  
summary(model)

airbag_mod <- glm(is_injured~ airbag + age_group, data = people, binomial(link = 'probit'))
summary(airbag_mod)

```

```{r results = 'asis'}
traffic_model <- traffic %>%
   mutate(is_severe = if_else(str_detect(crash_type, 'TOW'), TRUE, FALSE),
         four_way = if_else(roadtype == 'FOUR WAY', TRUE, FALSE),
         bad_weather = if_else(weather %in% c('RAIN', 'FREEZING RAIN/DRIZZLE', 'SLEET/HAIL', 'SNOW'), TRUE, FALSE),
         wet_road = if_else(str_detect(road_cond, 'SNOW | WET | ICE'), TRUE, FALSE)) %>%
  select(is_severe, four_way, bad_weather, wet_road)


model1 <- glm(is_severe~ four_way, data = traffic_model, binomial(link = 'probit'))

model2 <- glm(is_severe~ four_way + bad_weather, data = traffic_model, binomial(link = 'probit'))

model3 <- glm(is_severe~ four_way + bad_weather * wet_road, data = traffic_model, binomial(link = 'probit'))

stargazer::stargazer(model1, model2, model3, type = 'html', title = 'Chicago Crashes Model')
```

```{r}
Inputs {.sidebar}
-------------------------------------
```{r}
sliderInput("v_year", "Year:", 
            min = 2013, max = 2022, value = 2022, step = 1, sep = '')
```
```


``` {r}
# import data
traffic <- jsonlite::read_json('https://data.cityofchicago.org/resource/85ca-t3if.json', simplifyVector = T)
chi_sf <- st_read("https://data.cityofchicago.org/api/geospatial/5jrd-6zik?method=export&format=GeoJSON")

# check CRS
st_crs(traffic) # NA
st_crs(chi_sf)  # EPSG 4326 (WGS84)

# Convert lat/long from original data to sf
traffic_sf <- traffic %>% 
  as_tibble() %>%
  filter(lubridate::year(date_police_notified) == 2022) %>%
  drop_na(c(latitude, longitude)) %>%
  st_as_sf(coords = c('longitude', 'latitude'), crs = 4326)

# intersect points with polygons
sf_intersect <- st_intersection(traffic_sf, chi_sf)

# # join data of intersected points to polygon
sf_join <- left_join(chi_sf, st_drop_geometry(sf_intersect))
# calculate number of accidents per geoid
sf_group <- sf_join %>%
  group_by(geoid10) %>%
  summarise(accidents = n())

bins <- c(0, 2, 4, 6, 8, Inf)

sf_group %>%
  transform(accidents_bin = cut(accidents, breaks = bins)) %>%
  ggplot(aes(fill = accidents_bin)) +
  geom_sf(color = alpha("grey40", 0.5)) +
  scale_fill_brewer(palette = "GnBu") +
  theme_bw()
```